### 2015-07-31 tests/fitness.tex
tex

\documentclass[twocolumn]{article}
\usepackage{amsmath}
\begin{document}
\paragraph{Assessing the quality of reconstructed functions}
In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.

A measure that is timeshift independent and maximized for functions of identical shape is the 
maximum of the cross correlation 
\begin{align}
f_c(g, g^\prime) = \max_t \left<g, g^\prime(\cdot - t)\right> / \sqrt{\left< g, g\right> \left<g^\prime, g^\prime \right>}.
\end{align}
Algorithmically, the signals are split into chunks which are correlated separately. This saves computational power and 
prevents minute differences in frequency to accumulate into a huge phase shift, i.e. the phase only has to remain relatively
constant over the timescale of a single chunk.

\end{document}